---
title: "Assignment_b1_b2"
author: "Paul Li"
date: "2022-11-08"
output: github_document
---

## Setup
```{r}
library(tidyverse)
library(datateachr)
library(testthat)
```


## Exercise 1 & 2: Make a Function and Document Your Function

```{r}
#' Title Summary Statistics for factors of interest grouped by other factors of interest
#' Description Produce summary statistics on the factor of interest to be summarized on each group defined and specified by the grouping factor of interest
#' @param data is the dataset in use
#' @param group_factor is the variable/column to be grouped of interest in the data
#' @param summarize_factor is the variable/column to be summarized with different summary statistics
#'
#' @return the numeric minimum, first quartile (Q1), median, third quartile (Q3), maximum, range, and interquartile range (IQR), and standard deviation of the factor of interest to be summarized for all groups defined by the grouping factor of interest
#' @export
#'
#' @examples
#' neighbourhood_diameter_summary <- stat_summary(vancouver_trees, "neighbourhood_name", "diameter")
stat_summary <- function(data, group_factor, summarize_factor) {
  if(!is.character(data[[group_factor]]) || is.factor(data[[group_factor]])) {
    stop("I am so sorry, but this function only works for character or factor input for the factor to be grouped 
         \n", 
         "You have provided the group_factor object of class: ", class(data[[group_factor]])[1] 
         )
  }
  if (!is.numeric(data[[summarize_factor]])){
    stop("I am so sorry, but this function only works for numeric input for the factor to be summarized
         \n", 
         "You have provided the summarize_factor object of class: ", class(data[[summarize_factor]])[1]
         )
  }
  data %>% 
    group_by(.data[[group_factor]], na.rm=T) %>% 
    summarize(Mean = mean(.data[[summarize_factor]], na.rm=T),
              Minimum = min(.data[[summarize_factor]], na.rm=T),
              Q1 = quantile(.data[[summarize_factor]], 0.25, na.rm=T),
              Median = median(.data[[summarize_factor]], na.rm=T),
              Q3 = quantile(.data[[summarize_factor]], 0.75, na.rm=T),
              Maximum = max(.data[[summarize_factor]], na.rm=T),
              Range = range(.data[[summarize_factor]], na.rm=T),
              IQR = IQR(.data[[summarize_factor]], na.rm=T),
              Standard_deviation = sd(.data[[summarize_factor]], na.rm = T))}

```


## Exercise 3: Example

Working example producing the summary statistics for the `diameter` for all groups in `neighbourhood_name` (aka all neighbourhoods)
```{r, error = TRUE}
neighbourhood_diameter_summary <- stat_summary(vancouver_trees, "neighbourhood_name", "diameter")
```
Working example producing the summary statistics for the `diameter` in different `street_side_name`
```{r, error = TRUE}
street_side_diameter_summary <- stat_summary(vancouver_trees, "street_side_name", "diameter")
```

Failed example because of wrong input data type (input data type for `group_factor` placeholder is not character or factor; `summarize_factor` is still numeric or double)
```{r, error = TRUE}
## replace character or factor with date here
stat_summary(vancouver_trees, "date_planted", "diameter")

## replace character or factor with numeric (double) here
stat_summary(vancouver_trees, "height_range_id", "diameter")
```

Failed example because of wrong input data type (input data type for `summarize_factor` placeholder is not numeric; input data type for `group_factor` placeholder is still character or factor)
```{r, error = TRUE}
## replace numeric with character here
stat_summary(vancouver_trees, "curb", "neighbourhood_name")

## replace numeric with date here
stat_summary(vancouver_trees, "street_side_name", "date_planted")
```

Failed example with empty input specified for second placeholder (aka `group_factor`)
```{r, error = TRUE}
stat_summary(vancouver_trees, "", "diameter")
```

Failed example with empty input specified for third placeholder (aka `summarize_factor`)
```{r, error = TRUE}
stat_summary(vancouver_trees, "curb", "")
```

Failed example with wrong input column name specified
```{r, error = TRUE}
stat_summary(vancouver_trees, "curbs", "diameter")
```

Failed example without data specified (no data)
```{r, error = TRUE}
stat_summary("curb", "diameter")
```

## Exercise 4: Function Tests
```{r}

test_that("stat_summary tests", {
  expect_error(stat_summary(vancouver_trees, "", "diameter"))
  expect_error(stat_summary(vancouver_trees, "curb", ""))
  expect_error(stat_summary(vancouver_trees, "curb", "street_side_name"))
  expect_error(stat_summary(vancouver_trees, "diameter", "diameter"))
  expect_error(stat_summary("diameter", "diameter"))
  expect_error(stat_summary(vancouver_trees, "curbs", "diameter"))
})

```
